The Practical Application of Computerised Multistage Tests

### 2021-02-26 article.Rmd
Use ggplot for polynomial fit graph

---
title: "The Practical Application of Computerised Multistage Tests"
author:
- name: James Page
  affiliation: SP Educational Consulting
  affiliation_url: https://www.speduconsulting.co.uk/
date: "`r Sys.Date()`"
output:
  pdf_document:
    highlight: pygments
    latex_engine: lualatex
version: 1
description:
  A practical guide to the design and development of computerised multistage tests.
bibliography: references.bib
---

@@

@@

```{r Stage 1 MCC}
moduleA1_items <- which(colnames(dfTrial) %in% colnames(dfModuleAItems))
plotData <- trueScoreSummary(plot(raschmodel, 
                                  type = "ICC", 
                                  items = moduleA1_items, plot = FALSE), 
                             "A")
ggplot(plotData, aes(x=theta, y=trueScore)) + 
  geom_line(aes(colour=module)) +
  labs(title="Module Characteristic Curve", x="θ", y="True Score")
```
You can read this off such that a candidate who scores 9 in the module A section of the test would have an estimated ability score, ot $\hat{\theta}$ of approximately 2.

We can compare the MCCs for the stage 2 modules:

@@

ggplot(plotData, aes(x=theta, y=trueScore)) + 
  geom_line(aes(colour=module)) +
  labs(title="Module Characteristic Curve", x="θ", y="True Score")
```
A problem we can see already here is that these curves look too straight - ideally there should be a nice sigmoid shape to them. These MCCs have a low information content - in other words they are not very discriminating. We may want to move some items to provide a b etter shape to this curve.

@@

So, a candidate who scores 13 out of 20 on module A would have a $\hat\theta$ = `r gA1(13)`.

@@

```{r}
moduleA_TSRange <- data.frame(
  trueScore = c(seq(0,20)),
  theta = c(rep(0, 21))
)

# plot the theta/true score curve from the Rasch model
plot(x=df_moduleA_TS$theta, y=df_moduleA_TS$trueScore, 
     type="l", col = "coral1", main="Module A", xlab="θ", ylab="True Score")
moduleA_TSRange$theta <- gA1(moduleA_TSRange$trueScore)
points(x=moduleA_TSRange$theta, y=moduleA_TSRange$trueScore, col="purple", pch=4)
legend(2, 9, legend=c("MCC", "gA()"),
  col=c("coral1", "purple"), lty=1:2, cex=0.8)
```
We can see that it is a good fit, both in the middle of the range, and importantly at the upper and lower bounds.

@@

$$
s(\hat{\theta}_i) = k\frac{\hat{\theta}_i - min(\theta)}{max(\theta) - min(\theta)}
$$

@@

scaleScore <- s(thetaEstimate = -0.5, thetaRange = c(-4, 4))

@@

ggplot(plotData, aes(x=theta, y=information)) + 
  geom_line(aes(colour=module)) +
  labs(title="Module Information Function", x="θ", y="I(θ)") +
  geom_vline(xintercept = cutTheta, linetype="dotted")
```
Here we can see that there is point of intersection between modules B and C where $\hat\theta = $`r cutTheta`. If we read off the true score value that is equivalent for this theta estimate in Module A, then we can use this as the cut score. Below this value the candidate should be routed to Module B, above it they should be routed to Module C.

@@

8. Bibliography

>>>>

---
title: "The Practical Application of Computerised Multistage Tests"
author:
- name: James Page
  affiliation: SP Educational Consulting
  affiliation_url: https://www.speduconsulting.co.uk/
date: "`r Sys.Date()`"
output:
  pdf_document:
    highlight: pygments
    latex_engine: lualatex
  html_document:
    df_print: paged
version: 1
description: A practical guide to the design and development of computerised multistage
  tests.
bibliography: references.bib
---

@@

You can see that some of the outlier items have a very large divergence from the mean; it is best to remove these items from the item bank as they have too large an error to be used in a live test.

@@

```{r Stage 1 MCC}
moduleA1_items <- which(colnames(dfTrial) %in% colnames(dfModuleAItems))
plotData <- trueScoreSummary(plot(raschmodel, 
                                  type = "ICC", 
                                  items = moduleA1_items, plot = FALSE), 
                             "A")
ggplot(plotData, aes(x=theta, y=trueScore)) + 
  geom_line(aes(colour=module)) +
  labs(title="Module Characteristic Curve", x="θ", y="True Score")
```
You can read this off such that a candidate who scores 9 in module A would have an estimated ability score, or $\hat{\theta}$, of approximately 1.92.

We can go on to compare the MCCs for the stage 2 modules:

@@

ggplot(plotData, aes(x=theta, y=trueScore)) + 
  geom_line(aes(colour=module)) +
  labs(title="Module Characteristic Curve", x="θ", y="True Score")
```
A problem we can see already here is that these curves look too straight - ideally there should be a nice sigmoid shape to them. These MCCs have a low information content - in other words they are not very discriminating. We may want to move some items between modules, or reject items that have poor discrimination.

@@

So, a candidate who scores 9 out of 12 on module A would have a $\hat\theta$ = `r gA1(9)`.

@@

```{r}
df_gA <- data.frame(
  trueScore = c(seq(0,12)),
  theta = c(rep(0, 13))
)
df_gA$theta <- gA1(df_gA$trueScore)

# plot the theta/true score curve
ggplot(df_moduleA_TS, aes(x = theta, y = trueScore)) +
  geom_line(color = "red", linetype="dashed") +
  geom_point(data = df_gA, aes(x = theta, y = trueScore)) +
  scale_x_continuous(name="θ", limits=c(-4, 4)) +
  scale_y_continuous(name="True Score", limits=c(0, 12), 
                     minor_breaks = NULL, n.breaks = 12)
```
We can see that it is a good fit, both in the middle of the range, and importantly at the upper and lower bounds.

@@

$$
f(\hat{\theta}_i) = k\frac{\hat{\theta}_i - min(\theta)}{max(\theta) - min(\theta)}
$$

@@

scaleScore <- s(thetaEstimate = 0.95, thetaRange = c(-4, 4))

@@

ggplot(plotData, aes(x=theta, y=information)) + 
  geom_line(aes(colour=module)) +
  labs(title="Module Information Function", x="θ", y="I(θ)") +
  geom_vline(xintercept = cutTheta, linetype="dotted")
```
Here we can see that there is point of intersection between modules B and C where $\hat\theta = $ `r cutTheta`. If we read off the true score value that is equivalent for this theta estimate in Module A, then we can use this as the cut score. Below this value the candidate should be routed to Module B, above it they should be routed to Module C.

@@

8. References
